### 2017-01-30 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
Written up 1-3 of A1 in apx algs

\begin{document}
%Q1
\section{}
\paragraph{}
Let $E$ be the edge set of a graph $G$. Let $\cS$ be the collection of all sets $S \subseteq E$ that are matchings of $G$. Let $w: E \rightarrow \R_+$ be a weight function on the edges. For a subset $\bar{E} \subseteq E$ we denote by $G[\bar{E}]$ the subgraph of $G$ induced by the edges in $\bar{E}$.
	%Pa
\subsection{a}
\paragraph{}
Let $\bar{E} \subseteq E$. Suppose that $M \subseteq \bar{E}$ is an inclusion-wise maximal matching of $G[\bar{E}]$. Let $\bar{M}$ be a matching of $G[\bar{E}]$. Let $f : M \rightarrow \{ \{e_1,e_2\} | e_1, e_2 \in \bar{M}\}$ be defined by
$$ f(uv) = \{e_1, e_2\}$$
where $e_1 \in \bar{M}$ covers $u$ and $e_2 \in \bar{M}$ covers $v$, for each $uv \in M$.
\paragraph{}
We claim that for all $uv \in \bar{M}$, there exists $P \in Im(f)$ (where $Im(f)$ denotes the image of $f$) such that $uv \in P$. To see this suppose for a contradiction that there exists some $uv \in \bar{M}$ where each $P \in Im(f)$ does not contain $uv$. Then by the definition of $f$, neither $u$ nor $v$ are covered by $M$. But this violates that $M$ is inclusion-wise maximal since $uv \not\in M$ and $M \cup \{uv\}$ is a matching. Hence the claim holds.
\paragraph{}
From the previous claim we observe $|\bar{M}| \leq | \cup_{P \in Im(f)} P|$. But this yields a string of inequalities which complete the proof:
\begin{align*}
|\bar{M}| &\leq |\cup_{P \in Im(f)} P| \\
&\leq \sum_{P \in Im(f)} |P| \\
&\leq \sum_{P \in Im(f)} 2 \\
&\leq 2|Im(f)| \\
&\leq 2|M|. 
\end{align*}
With the last inequality following since $M$ is the domain of $f$. Hence we have $|M| \geq \frac{1}{2}|\bar{M}|$ as desired. $\blacksquare$
	%Pb
\subsection{b}
\paragraph{}
Let $e$

%Q2
\section{}

>>>>

\begin{document}
%Q1
Let $E$ be the edge set of a graph $G$. Let $\cS$ be the collection of all sets $S \subseteq E$ that are matchings of $G$. Let $w: E \rightarrow \R_+$ be a weight function on the edges. For a subset $\bar{E} \subseteq E$ we denote by $G[\bar{E}]$ the subgraph of $G$ induced by the edges in $\bar{E}$.
Let $\bar{E} \subseteq E$. Suppose that $M \subseteq \bar{E}$ is an inclusion-wise maximal matching of $G[\bar{E}]$. Let $\bar{M}$ be a matching of $G[\bar{E}]$. Let $f : M \rightarrow \{ \{e_1,e_2\} | e_1, e_2 \in \bar{M}\}$ be defined by
$$ f(uv) = \{e_1, e_2\}$$
where $e_1 \in \bar{M}$ covers $u$ and $e_2 \in \bar{M}$ covers $v$, for each $uv \in M$.
We claim that for all $uv \in \bar{M}$, there exists $P \in Im(f)$ (where $Im(f)$ denotes the image of $f$) such that $uv \in P$. To see this suppose for a contradiction that there exists some $uv \in \bar{M}$ where each $P \in Im(f)$ does not contain $uv$. Then by the definition of $f$, neither $u$ nor $v$ are covered by $M$. But this violates that $M$ is inclusion-wise maximal since $uv \not\in M$ and $M \cup \{uv\}$ is a matching. Hence the claim holds.
From the previous claim we observe $|\bar{M}| \leq | \cup_{P \in Im(f)} P|$. But this yields a string of inequalities which complete the proof:
|\bar{M}| &\leq |\cup_{P \in Im(f)} P| \\
&\leq \sum_{P \in Im(f)} |P| \\
&\leq \sum_{P \in Im(f)} 2 \\
&\leq 2|Im(f)| \\
&\leq 2|M|. 
With the last inequality following since $M$ is the domain of $f$. Hence we have $|M| \geq \frac{1}{2}|\bar{M}|$ as desired. $\blacksquare$
(This argument follows the exact same lines as the argument for problem $2$)
We can decided in polynomial time if $M\cup \{e\}$ is a matching in $\cS$ by iterating over $V$, counting the number of edges in $M\cup\{e\}$ which cover each vertex in $V$, and returning ``No" is the count is greater than $1$ for any vertex and ``Yes" otherwise. Thus the algorithm clearly runs in polynomial time and returns a feasible set.
Let $S$ be the set of edges returned by the greedy algorithm. Order the edges as $S = \{e_1, \dots, e_n\}$ in the order they were taken during the operation of $\cA$. Let $F_i := E\backslash \tilde{E}$ be the edges removed from $E$ at the edge of iteration $i$ of $\cA$ (the iteration where $e_i$ is added to $S$). Let $S^*$ be the optimal solution and partition $S^* = S^*_1 \cup \dots \cup S^*_n$ where $S^*_i = S^* \cap F_i$.
We want that $|S^*_i| \leq 2$ to hold, but we may need to redistribute some edges to ensure that this happens (and we will do this without effecting our ability to bound the optimal solution). Run the following algorithm, denoted $\cB$, to redistribute the edges in our partition of $S^*$:
\item For $i = 1, \dots, n$
\item While $|S^*_i| > 2$: take an edge $e$ from $S^*_i$ and redistribute $e$ to some $S^*_j$ for $j<i$ with $|S^*_j| < 2$.
\paragraph{Claim 1}
After the operation of $\cB$, for all $i = 1, \dots, n$, $|S^*_i| \leq 2$ and $w_{e_i} \geq w_e$ for all $e \in S^*_i$.
\paragraph{Proof of Claim 1}
We proceed by induction. For $e_1$ $\{e_1\}$ is inclusion-wise maximal on $\cS[F_1]$ (by definition of $F_1$), and $S_1^*$ is feasible on $\cS[F_1]$ since $S^*_1 \subseteq F_1$ and $S^* \in \cS$. Thus we have $|\{e_1\}| \geq \frac{1}{2} |S_1^*|$, and hence
$$|S^*_1| \leq \alpha.$$
Further by our greedy choice of $e_1$, $w_{e_1} \geq w_e$ for all $e \in F_1$ (and hence for all $e \in S^*_1$).
Suppose that the claim holds for $S^*_1 ,\dots, S^*_i$. Now by our greedy choice, and induction, $w_{e_{i+1}} \geq w_{e_j} \geq w_e$ for all $e \in S^*_j$ for all $j = 1,\dots, i$. So when $\cB$ redistributes edges of $S^*_{i+1}$ the cost property on the earlier sets is not violated. Also by our greedy choice, any edges that were not moved from $S^*_{i+1}$ by $\cB$ have cost at most $w_{e_{i+1}}$. To finish the proof of the claim, we demonstrate that while $|S_{i+1}^*| > \alpha$ there exists some $S^*_j$ with $j\leq i$ and $|S^*_j| < 2$. Observe that $\{e_1, \dots, e_{i+1}\}$ is inclusion-wise maximal on $\cS[F_1 \cup \dots \cup F_{i+1}]$ by definition of $F_1 \cup \dots \cup F_{i+1}$. Also $\bigcup_{j=1}^{i+1} S^*_j \in \cS[F_1 \cup \dots \cup F_{i+1}]$ so we have that
$$|\{e_1, \dots, e_{i+1}\}| \geq \frac{1}{2} |\bigcup_{j=1}^{i+1} S^*_j | = \frac{1}{2} \sum_{j=1}^{i+1} |S^*_j|.$$
Thus we see that
$$ \frac{1}{i+1}\sum_{j=1}^{i+1} |S^*_j| \leq 2.$$
Now since the desired bound holds on average, if $|S^*_{i+1}| > 2$ there exists some $S^*_j$ with $j \leq i$ and $|S^*_j| < 2$ (otherwise the bound does not hold). Thus we have proven Claim $1$. $\blacksquare$
With Claim $1$ in hand we are prepared to compute the desired approximation factor:
w(S^*) &= \sum_{i=1}^n c(S^*_i) \\
&\leq \sum_{i=1}^n w_{e_i} |S^*_i| \\
&\leq \sum_{i=1}^n w_{e_i} 2 \\
&= 2 w(S).
Thus $w(S) \geq \frac{1}{2} w(S^*)$ and hence the approximation factor holds as desired. $\blacksquare$
	%Pc
	\subsection{c}
	Suppose for a contradiction that the Greedy Algorithm is in fact a $(\frac{1}{2} + \epsilon)$-approximation algorithm for some $\epsilon \in (0, \frac{1}{2}]$. Then consider the following example. Let $G$ be the path graph with vertex set $V(G) = \{a,b,c,d\}$ and edge set $E(G) = \{ab, bc,cd\}$. Define the weight function $w : E(G) \rightarrow \R_+$ by
	$$w_{ab} = w_{cd} = 1 \quad\text{and}\quad w_{bc} = 1+\epsilon.$$
	Then the Greedy algorithm would choose the edge $bc$ as its solution. This solution has cost $1+\epsilon$. The optimal solution is to take edges $ab$ and $cd$ with cost $2$. The ratio of these solutions is
	$$ \frac{1+\epsilon}{2} = \frac{1}{2} + \frac{\epsilon}{2} < \frac{1}{2} + \epsilon$$
	Contradicting that the Greedy Algorithm is a $(\frac{1}{2} + \epsilon)$-approximation algorithm. $\blacksquare$
%Q2
Let $E$ be a set of $m$ elements and $\cS$ a collection of sets $S \subseteq E$. Let $w: E\rightarrow \R_+$ be a weight function. Let $\cS[\bar{E}] := \{ S \in \cS : S \subseteq \bar{E}\}$. Suppose that $\cS$ satisfies the following for a given $\alpha \geq 1$:
\item $S \in \cS$ implies $S' \in \cS$ for any $S' \subseteq S$
\item For all $\bar{E} \subseteq E$, if $S$ is an inclusion-wise maximal set in $\cS[\bar{E}]$ then $|S| \geq \frac{1}{\alpha} |\bar{S}|$ for all $\bar{S} \in \cS[\bar{E}]$.
We will show the Greedy Algorithm, denoted $\cA$, is a $\frac{1}{\alpha}$-approximation algorithm for the problem of finding a maximum weight $S \in \cS$.
Since we have a polynomial time oracle for membership in $\cS$ the algorithm clearly runs in polynomial time and returns a feasible set.
Let $S$ be the set of edges returned by the greedy algorithm. Order the edges as $S = \{e_1, \dots, e_n\}$ in the order they were taken during the operation of $\cA$. Let $F_i := E\backslash \tilde{E}$ be the edges removed from $E$ at the edge of iteration $i$ of $\cA$ (the iteration where $e_i$ is added to $S$). Let $S^*$ be the optimal solution and partition $S^* = S^*_1 \cup \dots \cup S^*_n$ where $S^*_i = S^* \cap F_i$.
We want that $|S^*_i| \leq \alpha$ to hold, but we may need to redistribute some edges to ensure that this happens (and we will do this without effecting our ability to bound the optimal solution). Run the following algorithm, denoted $\cB$, to redistribute the edges in our partition of $S^*$:
\item For $i = 1, \dots, n$
\item While $|S^*_i| > \alpha$: take an edge $e$ from $S^*_i$ and redistribute $e$ to some $S^*_j$ for $j<i$ with $|S^*_j| < \alpha$.
\paragraph{Claim 1}
After the operation of $\cB$, for all $i = 1, \dots, n$, $|S^*_i| \leq \alpha$ and $w_{e_i} \geq w_e$ for all $e \in S^*_i$.
\paragraph{Proof of Claim 1}
We proceed by induction. For $e_1$ $\{e_1\}$ is inclusion-wise maximal on $\cS[F_1]$ (by definition of $F_1$), and $S_1^*$ is feasible on $\cS[F_1]$ since $S^*_1 \subseteq F_1$ and $S^* \in \cS$. Thus we have $|\{e_1\}| \geq \frac{1}{\alpha} |S_1^*|$, and hence
$$|S^*_1| \leq \alpha.$$
Further by our greedy choice of $e_1$, $w_{e_1} \geq w_e$ for all $e \in F_1$ (and hence for all $e \in S^*_1$).
Suppose that the claim holds for $S^*_1 ,\dots, S^*_i$. Now by our greedy choice, and induction, $w_{e_{i+1}} \geq w_{e_j} \geq w_e$ for all $e \in S^*_j$ for all $j = 1,\dots, i$. So when $\cB$ redistributes edges of $S^*_{i+1}$ the cost property on the earlier sets is not violated. Also by our greedy choice, any edges that were not moved from $S^*_{i+1}$ by $\cB$ have cost at most $w_{e_{i+1}}$. To finish the proof of the claim, we demonstrate that while $|S_{i+1}^*| > \alpha$ there exists some $S^*_j$ with $j\leq i$ and $|S^*_j| < \alpha$. Observe that $\{e_1, \dots, e_{i+1}\}$ is inclusion-wise maximal on $\cS[F_1 \cup \dots \cup F_{i+1}]$ by definition of $F_1 \cup \dots \cup F_{i+1}$. Also $\bigcup_{j=1}^{i+1} S^*_j \in \cS[F_1 \cup \dots \cup F_{i+1}]$ so we have that
$$|\{e_1, \dots, e_{i+1}\}| \geq \frac{1}{\alpha} |\bigcup_{j=1}^{i+1} S^*_j | = \frac{1}{\alpha} \sum_{j=1}^{i+1} |S^*_j|.$$
Thus we see that
$$ \frac{1}{i+1}\sum_{j=1}^{i+1} |S^*_j| \leq \alpha.$$
Now since the desired bound holds on average, if $|S^*_{i+1}| > \alpha$ there exists some $S^*_j$ with $j \leq i$ and $|S^*_j| < \alpha$ (otherwise the bound does not hold). Thus we have proven Claim $1$. $\blacksquare$
With Claim $1$ in hand we are prepared to compute the desired approximation factor:
w(S^*) &= \sum_{i=1}^n c(S^*_i) \\
&\leq \sum_{i=1}^n w_{e_i} |S^*_i| \\
&\leq \sum_{i=1}^n w_{e_i} \alpha \\
&= \alpha w(S).
Thus $w(S) \geq \frac{1}{\alpha} w(S^*)$ and hence the approximation factor holds as desired. $\blacksquare$
